---
title: "Mini-Project 2"
author: "Ran Tao, Angela Huang, David Munoz Tord"
date: "October 31, 2017"
output: html_document
---


## Data Wrangling


This project is divided into two parts. In part1, we aim to figure out how the contribution amount from the committees  is distributed according to different organization types in Massachussates. In part2, we generalize our analysis to all individual states that the dataset contains. We combine our own data frames from getting variables mainly from raw data sets "individuals" and "committees". There are five organization types denoted as five letters in the raw data sets and we derive their meanings from original data set dictionary and rename them into plain English for easier understanding. 


```{r, message=FALSE, warning=FALSE, echo=FALSE}
library(tidyverse)
library(scales)
library(ggplot2)
load("house_elections.rda")
load("candidates.rda")
load("committees.rda")
load("contributions.rda")
load("individuals.rda")
```


```{r}
#selecting variables from raw datasets
indiv <- individuals %>% select(cmte_id, transaction_amt, occupation, employer, state) %>% filter(state=="MA")
commit <- committees %>% select(cmte_id, cmte_name, cmte_type, cmte_party_affiliation, org_type)

#recoding org_type
commit$org_type[commit$org_type=="C"] <- "Corporation"
commit$org_type[commit$org_type=="L"] <- "Labor organization"
commit$org_type[commit$org_type=="M"] <- "Membership organization"
commit$org_type[commit$org_type=="T"] <- "Trade association"
commit$org_type[commit$org_type=="V"] <- "Cooperative"
commit$org_type[commit$org_type=="W"] <- "Corporation w/o capital stock"
commit$org_type[commit$org_type==""] <- "NA"

#join indiv$commit and filter out the NA
df1 <- indiv %>% left_join(commit, by ="cmte_id")
df <- df1 %>% group_by(org_type) %>% summarise(summ=sum(transaction_amt)) %>% filter(org_type %in% c("Corporation", "Labor organization", "Membership organization", "Trade association", "Cooperative", "Corporation w/o capital stock"))

#plot1
ggplot(df, aes(x=factor(org_type), y=summ)) + geom_bar(stat="identity", aes(fill=factor(org_type))) + theme(axis.text.x=element_text(angle=90, hjust=1))+labs(x="Org Type", y="Sum Amount")+scale_y_continuous(breaks=c(0,1000000,2000000,2700000), label=c("0","1000000","2000000","2700000"))
```


From this graph, the y axis stands for the sum of the contribution amount. We can see from the graph that membership orgnization contributes the most, compared to other organization types in Massachusates, while Corporation without capital stock contributes the least. 


```{r fig.height=20, fig.width=22}
#selecting variables from raw datasets
indiv_state <- individuals %>% select(cmte_id, transaction_amt, occupation, employer, state)

#recoding $state
indiv_state$state[indiv_state$state==""] <- "NA"

#join indiv_state&commit and filter out the NA
df_state <- indiv_state %>% left_join(commit, by ="cmte_id")
df_state <- df_state %>% group_by(state, org_type) %>% summarise(summ=sum(transaction_amt)) %>% filter(org_type != "NA", state %in% c("AL","AK","AZ","AR","CA","CO","CT","DE","FL","GA","HI","ID","IL","IN","IA","KS","KY","LA","ME","MD","MA","MI","MN","MS","MO","MT","NE","NV","NH","NJ","NM","NY","NC","ND","OH","OK","OR","PA","RI","SC","SD","TN","TX","UT","VT","VA","WA","WV","WI","WY"))

#plot_state
ggplot(df_state, aes(x=factor(org_type), y=summ)) + geom_bar(stat="identity", position = position_dodge(), aes(fill=factor(org_type))) + theme(axis.text.x=element_text(angle=90, hjust=1), axis.title=element_text(size=15,face="bold"))+labs(x="Org Type", y="Sum Amount", size =5) + scale_y_continuous(breaks=c(0,1000000,3000000,5000000,7000000,9000000,11000000), label=c("0","1000000","3000000","5000000","7000000","9000000","11000000")) + facet_wrap(~state) 


```

This graph generalizes our anlaysis in MA to all individual states in the raw data sets.There are 50 states in total. The y axis represents the sum amount of the contribution from committees and the x axis of individual graph accounts for the five organization types. The whole graph aims to explain the distribution of contribution amount according to five organization types in 50 states. 

